---
title: "Try your data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.show = "animate",
                      fig.width = 9, fig.height = 7)

library(tidyverse)
library(gganimate)
source("helper_functions.R")
```

## 1. Prepare your data

### a) Initialize dataset as tibble of two variables
```{r}
set.seed(1234)
x <- rnorm(12, mean = rep(1:3, each = 4), sd = 0.8)
y <- rnorm(12, mean = rep(c(1, 2, 1), each = 4), sd = 0.8)
df_xy <- tibble(x, y)
```
### b) Set initial coordinates of cluster`s centers and their identificators
```{r}
Cx = c(1.6, 1.6, 2.5)
Cy = c(1, 2, 1.5)
# Number of clusters define as length of Cx or Cy vectors
n_clust <- length(Cx)
df_cl <- tibble(Cluster = factor(seq(n_clust), ordered = T), Cx, Cy)
```
### c) Make dataset that connect our points with initial cluster's
```{r}
df_kmeans0 <- bind_cols(df_xy, map_dfr(df_cl, rep, each = length(x) / n_clust))
```
### d) Now, take a look at the initial configuration of the clusters
```{r}
ggplot(df_kmeans0, aes(x, y)) +
  geom_point(shape = 1, size = 5) +
  geom_text(aes(x, y, label = rownames(df_xy)), vjust = 1.8, size = 4, color = "black") + 
  geom_point(aes(Cx, Cy, color = Cluster), shape = 3, size = 5) +
  labs(title = "The initial location of the cluster centers") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```
## 2. Run k-means algorithm

### a) Get dataset with k-means iterations
```{r}
n_iter <- 5 # nubmer of k-means iterations
df_kmeans <- wraper_kmeans(df_xy, df_cl, n_iter)
df_kmeans[[n_iter]]
```
### b) Compare result with build-in kmeans() 
```{r}
mxy <- as.matrix(df_xy)
mcl <- matrix(Cx, Cy, nrow = n_clust, ncol = dim(df_xy)[2])
test_kmeans <- kmeans(mxy, mcl, iter.max = n_iter)
test_kmeans
```

## 3. Visualize results

### a) Make "long" dataframe from the df_kmeans list
```{r}
df_long <- map_dfr(df_kmeans, ~ .)
```
### b) Animate points accesory to clusters (by changing color) and centers of clusters
```{r}
ggplot(df_long, aes(x, y, color = Cluster, frame = Iteration)) +
  geom_point(aes(group = Cluster), size = 5) +
  geom_point(aes(x = Cx, y = Cy), shape = 3, size = 7, color = "black") +
  geom_path(aes(x = Cx, y = Cy, cumulative = TRUE, group = Cluster),
            color = "black", size = 1) +
  geom_text(aes(x, y, label = rowid), vjust = 1.8, size = 4, color = 'black') +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) -> gg
gganimate(gg)
```
### c) Visualize clusters with poligons
```{r}
df_kmeans %>% 
  map(~ group_by(.x, Cluster) %>%
        by(.$Cluster, FUN = function(.) {.[chull(.$x, .$y), ]}) %>% 
        map_dfr( ~ .) %>% ungroup()) %>% 
  map_dfr( ~ .) -> df_polygon

gg_poly <- gg +
  geom_polygon(data = df_polygon,
               aes(x, y, group = Cluster, fill = Cluster, alpha = .5))
gganimate(gg_poly)
```

### d) Visualize of changing Euler distance for each iteration
```{r}
ggplot(df_long, aes(x = rowid, y = Eu_dist, frame = Iteration, cumulative = TRUE)) +
  geom_point(aes(color = Iteration), size = 5) +
  geom_line(aes(color = Iteration), size = 1) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) -> gg_dist
gganimate(gg_dist)
```
